The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-12-03 theoretical-consensus/article/article.tex
change hyphens
refactor on feedback

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

Following the line of the previous sections, it has been reasoned about the need 
for permissionless consensus when there are already well known and established consensus protocols
that work in trusted environments \cite{castro1999practical, miller2016honey}. However, even those protocols have
their own limitations, not only in terms of trust, fault-tolerance, centrality, permissions,
or bottlenecks, but also in terms of scalability \cite{miller2016honey}. Trying to put some effort on decentralization, 
Byzantine-Fault-Tolerant consensus protocols are not known for their performance when the network grows
in size. The more participants there are, the more messages need to be exchanged
between them, and the more time it takes to reach consensus, even if assuring deterministic finality \cite{decker2016bitcoin}. 
This is a problem that is not only related to the number of participants, but also to the
communication fashion and bandwidth, and the computational capacity of the devices that
participate in the consensus protocol. 
Summing up, the need for permissionless consensus is then justified by the fact that
permissioned protocols are not compatible with the requirements of the new generation of
distributed systems, especially in the context of blockchain networks. 
These requirements include dealing with today's sparse networks of anonymously and dynamically
participating devices, without interrupting consensus and while battling Sybil attacks \cite{8629877, survey-dist-consensus}.
Fundamentally, the permissionless consensus problem is the need for a consensus protocol that
can be run in a distributed and decentralized environment, where the participants are unknown and untrusted,
and where the network is bigger, sparser and unpredictably less reliable.

@@

There are a couple of particularities that need to be mentioned. First, the block generation
is of probabilistic nature, as the nonce is a random value, and the hash function is a one-way function.
Allied to this, increasing the hashing power is not an easy task, which consequently allows for tackling Sybil attacks,
despite the permissionless and pseudonymity nature of the network. Plus, the incentive mechanism
also plays a role in this regard, by naturally incentivizing honest participation.
The block generation interval, forged by the adjustment of the difficulty
threshold, is also of major importance because it reduces the fork probability and
allows for the timely propagation of proposed blocks, guaranteeing consensus
in the presence of an honest majority of the hashing power \cite{garay2015bitcoin, natoli2019deconstructing}.
This probabilistic finality shifts the 1/3 BFT threshold to a 1/2 threshold.

@@

These aim at achieving deterministic finality, by employing a BFT
consensus protocol, not at the block proposal level, but at the
block acceptance level. The idea is to have a free flow of blocks
proposed by the network, but to have a quorum of entities that
are responsible for the block acceptance. The block generation and proposal mechanism
can be of any kind, for example, Chain-based \emph{PoS}, or Committee-based \emph{PoS},
or even pure Proof-of-Work. The block acceptance layer then allows
for a deterministic acceptance of the blocks, by having a sort of permissioned
BFT consensus protocol, with the trivial 1/3 Byzantine tolerance. The most well known examples
and variations are the Tendermint \cite{buchman2016tendermint} and the new Ethereum Casper 
\cite{buterin2017casper} protocols.

@@

Permissionless consensus is a promising solution for these challenges, as an enabler for
decentralized environments. The use cases for these networks are diverse. Nevertheless,
as long as consensus is achieved, the network as a whole can be abstractly seen and used
as a trusted entity that runs as a trustless decentralized service. Applications may then
be built on top of it, and the network may be used as a platform for other services \cite{queralta2021blockchain}.
However, there are some key aspects that need to be addressed
in order to make this approach viable for resource-constrained networks, as we saw in the previous section:

@@

In this paper, it was presented a comprehensive summary of the permissionless
consensus problem, along with a short analysis of the most widely used consensus 
protocols, and the challenges that arise when applying these protocols to different environments.
In particular, this paper presented the challenges that resource-constrained networks may be facing, 
and the limitations of the existing consensus protocols when applied to this type of networks.
We have also presented the main approaches that have been explored to address these challenges,
and showed some potential paths for future research.

>>>>

% correct bad hyphenation here
\hyphenation{transi-tioning computa-tionally reaching tole-rance parti-cipants extrapo-late determi-nistic approach approaches issues capabi-lities limi-tations commu-nication}

@@

Following the line of the previous sections, it has been reasoned about the need 
for permissionless consensus when there are already well known and established consensus protocols
that work in trusted environments \cite{castro1999practical, miller2016honey}. However, even those protocols have
their own limitations, not only in terms of trust, fault-tolerance, centrality, permissions,
or bottlenecks, but also in terms of scalability \cite{miller2016honey}. Trying to put some effort on decentralization, 
Byzantine-Fault-Tolerant consensus protocols, hereinafter abbreviated as BFT protocols, 
are not known for their performance when the network grows
in size. The more participants there are, the more messages need to be exchanged
between them, and the more time it takes to reach consensus, even if assuring deterministic finality \cite{decker2016bitcoin}. 
This is a problem that is not only related to the number of participants, but also to the
communication fashion and bandwidth, and the computational capacity of the devices that
participate in the consensus protocol. 
Summing up, the need for permissionless consensus is then justified by the fact that
permissioned protocols are not compatible with the requirements of the new generation of
distributed systems, especially in the context of blockchain networks. 
These requirements include dealing with today's sparse networks of anonymously and dynamically
participating devices, without interrupting consensus and while battling Sybil attacks \cite{8629877, survey-dist-consensus}.
Fundamentally, the permissionless consensus problem is the need for a consensus protocol that
can be run in a distributed and decentralized environment, where the participants are unknown and untrusted,
and where the network is bigger, sparser and unpredictably less reliable.

@@

There are a couple of particularities that need to be mentioned. First, the block generation
is of probabilistic nature, as the nonce is a random value, often of the form of a counter, 
and the hash function is a one-way function.
Allied to this, increasing the hashing power is not an easy task, which consequently allows for tackling Sybil attacks,
despite the permissionless and pseudonymity nature of the network. Plus, the incentive mechanism
also plays a role in this regard, by naturally incentivizing honest participation, at the application level, 
for example, in the form of monetary rewards, in the case of Bitcoin \cite{nakamoto2008bitcoin}.
The block generation interval, forged by the adjustment of the difficulty
threshold, is also of major importance because it reduces the fork probability and
allows for the timely propagation of proposed blocks, guaranteeing consensus
in the presence of an honest majority of the hashing power \cite{garay2015bitcoin, natoli2019deconstructing}.
This probabilistic finality shifts the 1/3 BFT threshold to a 1/2 threshold.

@@

These aim at achieving deterministic finality, by employing a BFT
consensus protocol, not at the block proposal level, but at the
block validation or acceptance level (See Fig. \ref{fig:building-blocks-consensus}).
The idea is to have a free flow of blocks
proposed by the network, but to have a quorum of entities that
are responsible for the block acceptance. The block generation and proposal mechanism
can be of any kind, for example, Chain-based \emph{PoS}, or Committee-based \emph{PoS},
or even pure Proof-of-Work. The block acceptance layer then allows
for a deterministic acceptance of the blocks, by having a sort of permissioned
BFT consensus protocol, with the trivial 1/3 Byzantine tolerance. The most well known examples
and variations are the Tendermint \cite{buchman2016tendermint} and the new Ethereum Casper 
\cite{buterin2017casper} protocols.

@@

Permissionless consensus is a promising solution for these challenges, as an enabler for
decentralized environments. The use cases for these networks are diverse. Nevertheless,
as long as consensus is achieved, the network as a whole can be abstractly seen and used
as a trusted entity that runs as a trustless decentralized service. Applications may then
be built on top of it, and the network may be used as a platform for other services \cite{queralta2021blockchain}.
However, there are some key aspects that need to be addressed
in order to make this approach viable for resource-constrained networks, with regard to their characteristics. These aspects are:

@@

In this paper, it was presented a comprehensive summary of the permissionless
consensus problem, along with a short analysis of the most widely used consensus 
protocols, and the challenges that arise when applying these protocols to different environments.
We reasoned about the need for permissionless consensus, and showed that most of the existing
consensus protocols follow a very similar phased design, often differing only in the block
generation phase, or in the consensus proof units, concluding that those design choices are usually 
related with the trade-offs imposed by the blockchain trilemma.
We also presented some other approaches that deviate from that common
structural design, and that may be more suitable for some specific network environments.
In particular, this paper finished presenting the challenges that resource-constrained networks may be facing, 
and the limitations of the existing consensus protocols when applied to this type of networks.
We also argued about the main approaches that have been explored to address these challenges,
and showed some potential paths for future research.
